# Data

### About Delta Science Tracker  

We are using publicly available data from the Delta Stewardship Council's (DSC) [Delta Science Tracker](https://sciencetracker.deltacouncil.ca.gov/). This database is a library of Sacramento and SJ Delta-related research over the last 50 years. As the DSC's website explains:

| "The Delta Science Tracker is a tool to improve coordination and collaboration of science activities in a way that is valuable to scientists, decision-makers, and managers in the Delta. The Delta Science Tracker allows users to explore and visualize recent research and monitoring efforts through the lenses of science activities, organizations, and people. It is intended to promote communication, create opportunities for collaboration, and enhance transparency for science funding opportunities and decisions."

The data we'll use was downloaded from the DSC's [visualizations page](https://sciencetracker.deltacouncil.ca.gov/visualizations), with the collaborations by organizations filter. Additional data was downloaded through scraping their webpage, which we did with permission from the DSC. Code for scraping their webpage can be found [here](https://github.com/ucdavisdatalab/workshop_network_viz/blob/main/scraping_data.R).

### Data descriptions  

Before we get into visualizations, we want to take a look at the data we have available. What relationships do we have? And what node or edge attributes are available? 

**Two-mode edge list:** The data was originally compiled as two-mode network data: how are organizations/researchers connected to research projects? Our first mode, `project_id` are the id numbers representing different Delta science research projects. And our second mode, `org_id`, is the organization(s) (agencies, universities, etc) that are working on that project. This two-mode edge list has 680 rows, whereby 138 organizations are connected with one or more of 297 projects. In these data, organizations often take part in more than one project, and projects often have more than one organizations associated with it. We also have edge attributes, which represent the kind of involvement an organization has in the project: contributor or leader.  

```{r}
edges <- read.csv('data/edgelist_twomode.csv')
head(edges)
```

```{r, eval = F, echo = F}
length(unique(edges$org_id))
length(unique(edges$project_id))
```

**One-mode edge list:** We projected our two-mode network to a one-mode network to take a deeper look at research collaborations: What organizations are working together? We've also added in an edge attribute to our collaboration tie based on the years that different actors collaborated together.

```{r}
edges_p <- read.csv('data/edgelist_projected.csv') 
head(edges_p)
```

```{r, echo = F}
# We got out projected data directly from the DST webpage, but generally,
# you can project any bipartite network into a one-mode network using the
# following steps:

# create count table where organizations are columns (make y)
tbl <- table(edges[c(2,1)])
# pull out column names -- these are mode 1 (orgs)
orgids <- colnames(tbl)
# take the cross-product to get a co-occur matrix
comat <- crossprod(tbl)
# assign self-co-occurrences 0
diag(comat) <- 0
# This should be symmetrical
isSymmetric(comat)
# make a df and add ids as the column names
comat <- data.frame(comat)
colnames(comat) <- orgids
dim(comat)
# Remove those with no co-occurrence at all
comat2 <- comat[rowSums(comat, na.rm = T) != 0, 
                colSums(comat, na.rm = T) != 0]
dim(comat2)
# make co-occur df into matrix object
comat3 <- as.matrix(comat2)

# create a graph so that it can be converted into a weighted edgelist
g <- igraph::graph_from_adjacency_matrix(comat3, weighted = T, mode = 'undirected')
el_proj <- igraph::get.data.frame(g)

# rename columns and sort
colnames(el_proj) <- c("V1", "V2", "weight")
```

**Nodelist**: We have a 'master' list of all the nodes (organizations and projects), and some attribute information. We have more node attributes for projects than we do for organizations. For projects, we have funding data, the dates of the project, as well as a series of categorical tags related to what kinds of scientific and management themes the project addresses. For organizations, we have very few attributes, but for the sake of this presentation we projected the scientific and management theme attributes from projects onto organizations that worked on those projects.  

```{r}
nodes <- read.csv('data/nodelist_twomode.csv')
head(nodes)
```
Note that the scientific and management themes are formatted as Boolean values for each theme. This is because projects can have more than one theme, and so listing them out would cause our data to be 'long' with repeated project observations. With node lists, we want to only have unique observations. If your data are 'long' (with repeating observations), widen them out as we did with the scientific and management themes. Further note: Formatting network data is half the battle. See our script if you want to see how we 

### What story do we want to tell?

One-mode:
* How have different organizations collaborated across projects?
* Do we see any particular research communities emerge across these collaborations? (community detection + representing node attributes by color)
* How have those collaborations changed over time? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)

Two-mode: 
* What actors have been most active across projects? 
* Are most-active members typically leaders or contributors? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)
* What actors have been most active on climate-related projects? (Node attribute of project node)



