# Creating network objects in R

Now that we have our edge lists and node lists, we want to create our network objects. There are two popular network packages (and consequently, network object types) in R. `igraph` generates a list object of class "igraph" and `network` generates an list object of class "network". We are going to be working primarily with `network` objects, but I'd like to start by making both because you are likely to run into both of them along the way, and both are compatible with the `ggraph` package that we will be using. 

```{r setup, include=FALSE}
knitr::opts_knit$set(warning = F, message = F)
```

First, let's read in our two network packages, as well as a couple other packages that we'll need for this workshop.

```{r, results = F, message=F, warning = F}
library(igraph)
library(network)
library(dplyr)
library(magrittr)
```


```{r, echo = F}
edges_2mode <- read.csv('data/edgelist_twomode.csv')
edges_1mode <- read.csv('data/edgelist_projected.csv')
nodes_2mode <- read.csv('data/nodelist_twomode.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')
```

## `igraph` 

First, let's make an `igraph` object. `igraph` generally seems to have more documentation/Stack Overflow resources than other network packages, so it might be tempting to start here. We can load in the `igraph` package and use the `graph_from_data_frame()` function to create our graph. Note that there are dozens of `graph_from...` functions. Many of them sound (and are) intuitive. When dealing with edge and node lists, I prefer `graph_from_data_frame()` over `graph_from_edgelist()` because the former allows us to include the node list as an argument, whereas the latter does not.

We want to start with creating both our one and two-mode networks. For our one-mode networks we want to use the `edges_1mode` and `nodes_1mode` objects. 

```{r}
g1 <- graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

We can take a look at a summary of the igraph object below. We see the number of nodes (139) and edges (475) summarized at top and our list of node attributes (followed by (v/[datatype])) and edges attributes (followed by (e/[datatype])).

```{r}
summary(g1)
```

Two-mode networks are a bit harder to create in `igraph`. Generally this package has very little two-mode functionality. There is one function for this, `make_bipartite_graph()`, but it requires different inputs than the standard edge list or matrix, and node lists. Instead, this function requires a vector of vertex types, meaning a Boolean of modes (0-1), and a vector of the edges in the graph. 

```{r, eval= F}
?make_bipartite_graph()
```

We won't walk through the steps on how to do this, but you can reveal the steps below if you want to know more.  

<details>
<summary>**Reveal:** How to make a two-mode network from edge list and node list in igraph</summary>
```{r}
# edges need to be a as a vector where each pair is a connection, 
# so we can vectorize the edge list (e.g. convert the connection between: 
# id 49590 and id 50122 be just paired together as our first two values 
# in the vector, and so on
edges_2mode_v <- as.vector(matrix(rbind(edges_2mode$org_id,
                                        edges_2mode$project_id),ncol=1))
# But edges ALSO need to be numbered in a way that is compatible with the nodelist
nodes_2mode$id_numeric <- 1:nrow(nodes_2mode)

# So now we can reassign the edge list
library(dplyr)
library(magrittr)
edges_2mode <- left_join(edges_2mode, select(nodes_2mode, id, id_numeric), 
                         by = c('org_id' = 'id')) %>% 
  rename('org_id_numeric' = 'id_numeric') %>% 
  left_join(select(nodes_2mode, id, id_numeric), 
                         by = c('project_id' = 'id')) %>% 
  rename('project_id_numeric' = 'id_numeric')

# And re-make this vector
edges_2mode_v <- as.vector(matrix(rbind(edges_2mode$org_id_numeric,
                                        edges_2mode$project_id_numeric),ncol=1))

# Now we can make the two mode network in igraph
g2 <- make_bipartite_graph(nodes_2mode$mode, edges_2mode_v, directed = FALSE)
# And we have to manually assign attributes
g2 <- set_vertex_attr(g2, "id", value = nodes_2mode$id)
g2 <- set_vertex_attr(g2, "n_years", value = nodes_2mode$n_years)
g2 <- set_edge_attr(g2, 'leadership', value = edges_2mode$leadership)
g2 <- set_edge_attr(g2, 'contribution', value = edges_2mode$contribution)
```
</details>
<br>

Still, we can take a look. Much like the one-mode graph, we see the number of nodes (now 435) and edges (680) summarized at top and our list of node attributes and edges attributes.  

```{r}
g2
```

The syntax for manipulating igraphs is demonstrated below. We use the `V()` and `E()` functions to call on either vertices or edges of our igraph, respective. hen can use the `$` symbol to subset from those elements of the network.  

```{r}
# Subset the vertex attribute: the names of the vertices
head(V(g1)$name)
```

```{r}
# Subset the edge attribute: a logical value for whether a collaboration
# happened between 1995-2009
head(E(g1)$Y1995_2009)
```

## `network`

Personally, I find that the `igraph` package has more limited statistical functionality than `network` package (and its associated `statnet` suite of packes), so I prefer `network` objects. We initialize a network object with the `network` function (`network` really needs to get more creative in its naming approaches). Just like with `igraph`, for our one-mode network we will use our projected edges and mode one nodes. For our two-mode network we can use our full edge and node lists.

Note: You'll often see folks converting igraphs to network objects using the `intergraph` package. 

```{r}
library(intergraph)
net1_convert <- asNetwork(g1)
net1_convert
```

This conversion function can be very useful, but be sure to double check your work. Right now, with a relatively simple network (undirected and unweighted) the conversion works fine, but in other cases the default settings may require some attention. Because of this, personally I like to generate network objects directly from my edge and node lists.  

Within the `network` package we can use the `network` function and read in our edge lists (as the x argument), vertices, and specify features of the network such as whether or not it is directed or bipartite. 

```{r}
net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)

net2 <- network(x = edges_2mode, 
               vertices = nodes_2mode,
               bipartite = T,  
               directed = F)
```

When we print out a network object we get a different-looking summary, but it generally carries the same information: 
```{r}
net1
```

```{r}
net2
```

The syntax for manipulating `network` objects is demonstrated below. We use the `%v%` and `%e%` operators to call on either vertices or edges of our network, and then name the attribute in quotation marks 

```{r}
tail(net2 %v% 'n_years')
```

```{r}
head(net1 %e% 'Y1995_2009')
```

To save yourself trouble in the long run, it may be best to consider your network type, what kinds of analyses and visualization you'd like to do, and select a network package to stick to for that. Because there are differences in how these objects interact with various network analytical functions and  visualization tools, and how they are indexed (`V()$` vs `%v%`), it is worthwhile to stay consistent.  



