# Creating network objects in R

Now that we have our edge lists and node lists, we want to create our network objects. There are two popular network packages (and consequently, network object types) in R `igraph` generates a list object of class "igraph" and `network` generates an list object of class "network". We are going to be working primarily with `network` objects, but I'd like to make both because you are likely to run into both of them along the way. 

```{r, setup, include=FALSE}
knitr::opts_knit$set(warning = F, message = F)
```

First, let's read in our two network packages, as well as a couple other packages that we'll need for this workshop.

```{r}
library(dplyr)
library(magrittr)
library(igraph)
library(network)
library(sna)
library(ggraph)
```


```{r, echo = F}
edges_2mode <- read.csv('data/edgelist_twomode.csv')
edges_1mode <- read.csv('data/edgelist_projected.csv')
nodes_2mode <- read.csv('data/nodelist_twomode.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')
```

## `igraph` 

First, let's make an `igraph` object. `igraph` generally seems to have more documentation/Stack Overflow resources than other network packages, so it might be tempting to start here. We can load in the `igraph` package and use the `graph_from_data_frame()` function to create our graph. Note that there are dozens of `graph_from...` functions. Many of them sound (and are) intuitive. When dealing with edge and node lists, I prefer `graph_from_data_frame()` over `graph_from_edgelist()` because the former allows us to include the node list as an argument, whereas the latter does not.

We want to start with creating both our one and two-mode networks. For our one-mode networks we want to use the `edges_1mode` and `nodes_1mode` objects. 

```{r}
g1 <- graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

We can take a look at a summary of the igraph object below. We see the number of nodes (139) and edges (475) summarized at top and our list of node attributes (followed by (v/[datatype])) and edges attributes (followed by (e/[datatype])).

```{r}
summary(g1)
```

Two-mode networks are a bit harder to create in `igraph`. Generally this package has very little two-mode functionality. There is one functio fro this, `make_bipartite_graph()`, but it requires different inputs than the standard edge list or matrix, and node lists. Instead, this function requires a vector of vertex types, meaning a Boolean of modes (0-1), and a vector of the edges in the graph. 

```{r, eval= F}
?make_bipartite_graph()
```


```{r, echo = F}
# edges need to be a as a vector where each pair is a connection, 
# so we an vectorize to make from: 49590, to: 50122 be just paired together
# as our first two values in the vector, and so on
edges_2mode_v <- as.vector(matrix(rbind(edges_2mode$org_id,
                                        edges_2mode$project_id),ncol=1))
# But edges ALSO need to be numbered in a way that is compatible with the nodelist
nodes_2mode$id_numeric <- 1:nrow(nodes_2mode)
# so reassign
edges_2mode <- left_join(edges_2mode, select(nodes_2mode, id, id_numeric), 
                         by = c('org_id' = 'id')) %>% 
  rename('org_id_numeric' = 'id_numeric') %>% 
  left_join(select(nodes_2mode, id, id_numeric), 
                         by = c('project_id' = 'id')) %>% 
  rename('project_id_numeric' = 'id_numeric')

# So we re-make this vector
edges_2mode_v <- as.vector(matrix(rbind(edges_2mode$org_id_numeric,
                                        edges_2mode$project_id_numeric),ncol=1))

g2 <- make_bipartite_graph(nodes_2mode$mode, edges_2mode_v, directed = FALSE)
# then we have to manually assign attributes
g2 <- set_vertex_attr(g2, "id", value = nodes_2mode$id)
g2 <- set_vertex_attr(g2, "n_years", value = nodes_2mode$n_years)
g2 <- set_edge_attr(g2, 'leadership', value = edges_2mode$leadership)
g2 <- set_edge_attr(g2, 'contribution', value = edges_2mode$contribution)
```


```{r}
g2
```

The syntax for manipulating igraphs is below. We use the `V()` and `E()` functions to call on either vertices or edges of our igraph, and then can use the `$` symbol to subset from those elements of the network.  

```{r}
# Subset the vertex attribute: the website of these projects
head(V(g2)$n_years)
```

```{r}
# Subset the edge attribute: a logical value for whether a collaboration
# happened between 1995-2009
head(E(g1)$Y1995_2009)
```

## `network`

Personally, I think the `igraph` has more limited statistical functionality than `network`, so I prefer `network` objects. We initialize a network object with the `network` function (`network` really needs to get more creative in its naming approaches). Just like with `igraph`, for our one-mode network we will use our projected edges and mode 1 nodes. For our two-mode network we can use our full edge and node lists.

You'll often see folks converting igraphs to network objects using the `intergraph` package. 

```{r}
library(intergraph)
net1_convert <- asNetwork(g1)
net1_convert
```


```{r}
net2_convert <- asNetwork(g2)
net2_convert
```
But personally I like to make it on my own to be sure. There are some features that don't always carry over (weighted edges)

```{r}
library(network)

net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)

net2 <- network(x = edges_2mode, 
               vertices = nodes_2mode,
               bipartite = T,  
               directed = F)
```

When we print out a network object we get a different-looking summary, but it generally carries the same information: 
```{r}
net1
```

```{r}
net2
```

The syntax for manipulating `network` objects is below. We use the `%v%` and `%e%` WHT ARE THESE CALLED to call on either vertices or edges of our network, and then name the attribute in quotation marks 

```{r}
tail(net2 %v% 'n_years')
```

```{r}
head(net1 %e% 'Y1995_2009')
```


