# Creating network objects in R

Now that we have our edge lists and node lists, we want to create our first network object. We are going to start by making a one-mode network.  

There are two popular network packages (and consequently, network object types) in R. `igraph` generates a list object of class "igraph" and `network` generates an list object of class "network". We are going to be working primarily with `network` objects, but I'd like to start by making both because you are likely to run into both of them along the way, and both are compatible with the `ggraph` package that we will be using. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, fig.align = 'center')
```

So, let's read in our two network packages.

```{r, results = F, message=F, warning = F}
library(igraph)
library(network)
```


```{r, echo = F}
edges_1mode <- read.csv('data/edgelist_onemode_projected.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')
```

## `igraph` 

First, let's make an `igraph` object. `igraph` generally seems to have more documentation/Stack Overflow resources than other network packages, so it might be tempting to start here. We can load in the `igraph` package and use the `graph_from_data_frame()` function to create our graph. Note that there are dozens of `graph_from...` functions. Many of them sound (and are) intuitive. When dealing with edge and node lists, I prefer `graph_from_data_frame()` over `graph_from_edgelist()` because the former allows us to include the node list as an argument, whereas the latter does not.

For our one-mode networks we want to use the `edges_1mode` and `nodes_1mode` objects. 

```{r}
g1 <- graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

We can take a look at a summary of the igraph object below. We see the number of nodes (139) and edges (475) summarized at top and our list of node attributes (followed by (v/[datatype])) and edges attributes (followed by (e/[datatype])).

```{r}
summary(g1)
```

The syntax for manipulating igraphs is demonstrated below. We use the `V()` and `E()` functions to call on either vertices or edges of our igraph, respective. hen can use the `$` symbol to subset from those elements of the network.  

```{r}
# Subset the vertex attribute: the names of the vertices
head(V(g1)$name)
```

```{r}
# Subset the edge attribute: a logical value for whether a collaboration
# happened between 1995-2009
head(E(g1)$Y1995_2009)
```

## `network`

Personally, I find that the `igraph` package has more limited statistical functionality than `network` package (and its associated `statnet` suite of packes), so I prefer `network` objects. We initialize a network object with the `network` function (`network` really needs to get more creative in its naming approaches). Just like with `igraph`, for our one-mode network we will use our projected edges and mode one nodes. 

Note: You'll often see folks converting igraphs to network objects using the `intergraph` package. 

```{r}
library(intergraph)
net1_convert <- asNetwork(g1)
```

This conversion function can be very useful, but be sure to double check your work. Right now, with a relatively simple network (undirected and unweighted) the conversion works fine, but in other cases the default settings may require some attention. Because of this, personally I like to generate network objects directly from my edge and node lists.  

Within the `network` package we can use the `network` function and read in our edge lists (as the x argument), vertices, and specify features of the network such as whether or not it is directed or bipartite. 

```{r}
net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)
```

When we print out a network object we get a different-looking summary, but it generally carries the same information: 

```{r}
net1
```

The syntax for manipulating `network` objects is demonstrated below. First, we can use `network`'s series of `get.` functions to index components of the network. For node attributes, you can use `get.vertex.attribute`:

```{r}
head(get.vertex.attribute(net1, 'name'))
```

For edge attributes, you can use `get.edge.attribute`:

```{r}
head(get.edge.attribute(net1, 'before_1980'))
```

You can also use special operators, `%v%` and `%e%`,  call on either vertices or edges of our network and then name the attribute in quotation marks 

```{r}
head(net1 %v% 'name')
```

```{r}
head(net1 %e% 'Y1995_2009')
```


### Preparing to work with `network`  

For the remainder of this tutorial are going to stick to working with the `network` package, even though the functions we will use for visualization are compatible with both objects. I propose using network objects because they are compatible with more advanced statistical analysis provided through the `statnet` suite of packages. 

Because we will be using exclusively `network` objects, we want to detach `igraph` before we continue further. This is because there are several commonly-used network functions in both `igraph` and `sna` that mask one another. For instance:

```{r, eval = F}
sna::degree()
igraph::degree()
```

The `igraph` degree function will work only on `igraph` objects and `sna` `degree` function will work only on `network` objects. To avoid confusion and masking in R, we are going to detach the `igraph` package and work only with `network` objects and compatible packages like `sna`.  

```{r}
detach("package:igraph", unload = TRUE)
```


