# Plotting packages

```{r, setup, include=FALSE}
knitr::opts_knit$set(warning = F, message = F, fig.align = 'center')
```

```{r}
library(dplyr)
library(magrittr)
library(igraph)
library(network)
library(sna)
library(ggraph)
```

```{r, echo = F}
edges_1mode <- read.csv('data/edgelist_projected.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')

net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)
g1 <- igraph::graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

## `ggraph`  

`ggraph` follows the grammar of graphics ...

[explain ggraph basics]

`ggraph` uses the same approach as the `ggplot2` grammar of graphics. This means is has three core components: 1. (network) data, 2. geometries, in this case edges and nodes, and 3. aesthetic mappings


```{r}
ggraph(net1) +
    geom_edge_link() +
    geom_node_point()
```


## Other packages and functions  

There are several plotting packages out there, some of them working with different network objects and requiring different knowledge. Excellent overviews of some of them are provided in this [post by Katya Ognyanova](https://kateto.net/network-visualization). I won't need to repeat these, but it is worth seeing the basic representation.

### `plot()`

Base R plotting works with network and igraph objects, though their defaults for each object are different. 

```{r, out.width=c('50%', '50%')}
plot(net1)
plot(g1)
```

### `GGally::ggnet2()`

The `ggnet2` function is also dynamic, but despite being housed as part of a ggplot2 series it seems to function more like a base R plot
https://briatte.github.io/ggnet/
```{r, out.width=c('50%', '50%')}
GGally::ggnet2(net1)
GGally::ggnet2(g1)
```

### `ggnetwork`

Very close to `ggraph`: https://cran.r-project.org/web/packages/ggnetwork/vignettes/ggnetwork.html
