# One-mode visualization

## Visualization goals:

1. What does the collaborative network of Delta science research look like? 
2. Do we see any particular research communities emerge across these collaborations? (community detection + representing node attributes by color)
3. How have research collaborations changed over time? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)

### Preparing to work with `network`

We are going to stick to only working with the `network` package, even though the functions we will use for visualization are compatible with both objects. However, there are commonly-used network functions in each package, `igraph` and `network`, that mask one another. For instance:

```{r}
sna::degree()
igraph::degree()
```

To avoid confusion and masking in R, we are going to detach the igraph package.

```{r}
detach("package:igraph", unload = TRUE)
```

```{r, echo = F}
edges <- read.csv('data/edgelist_twomode.csv')
edges_p <- read.csv('data/edgelist_projected.csv')
nodes <- read.csv('data/nodelist_twomode.csv')
nodes_mode1 <- nodes[nodes$mode == 1 & 
                       (nodes$id %in% edges_p$from_org_id | 
                          nodes$id %in% edges_p$to_org_id),]
net <- network(x = edges_p, 
               vertices = nodes_mode1,
               bipartite = F,  
               directed = F)
```


```{r}
net %v% 'degree' <- sna::degree(net)
```

## Getting started with `ggraph`

To introduce the basics of the `ggraph` package, we're going to focus on our first question: _What does the collaborative network of Delta science research look like?_ 

`ggraph` uses the same approach as the `ggplot2` grammar of graphics. This means is has three core components: 1. (network) data, 2. geometries, in this case edges and nodes, and 3. aesthetic mappings

```{r}
ggraph(net) +
    geom_edge_link() +
    geom_node_point()
```

These are some defaults at play here:
* layout = 'auto', which is the 'stress' layout  
* default aesthetic mapping for geom_edges: x, y, xend, yend
* default aesthetic mapping for geom_node_point: x, y

First, let's talk about layouts.
```{r}
?ggnet2
# Not helpful
# link to the website
```

```{r}
ggraph(net, layout = 'stress', bbox = 5) +
    geom_edge_link() +
    geom_node_point()

ggraph(net, layout = 'kk') +
    geom_edge_link() +
    geom_node_point()

ggraph(net, layout = 'fr') +
    geom_edge_link() +
    geom_node_point()

ggraph(net, layout = 'tree') +
    geom_edge_link() +
    geom_node_point()

ggraph(net, layout = 'circle') +
    geom_edge_link() +
    geom_node_point()
```
### Aesthetics

Now let's start adding some aesthetics, much like we would with a ggplot to improve overall appeal

Reminder about colors. I liek to set a colors object and call upon it
```{r}
viridis::viridis(8)
```
```{r}
clrs <- c("#440154FF", "#46337EFF", "#365C8DFF", "#277F8EFF",
          "#1FA187FF", "#4AC16DFF", "#9FDA3AFF", "#FDE725FF")
```

We can change colors and assign additional aestehtics to the figure.
```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree'), color = clrs[4]) 
```

We can also add themes and labels, just like with ggplot2. 
```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree'), color = clrs[4]) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Delta Science Collaborative Research Network")
```

Notice it changes every time, a little bit. You can set your seed (every time before you plot to keep it consistent.)

We might want to remove our legend
```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree'), color = clrs[4]) +
  theme_void() +
  theme(legend.position = "none")
```

### Label particular nodes

We have a variable that is our names, but it might be a bit overwhelming
```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree'), color = clrs[4]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = net %v% 'name'),
                family = "Times",
                size = 3, 
                color="black")
```

Let's be selective based on degree. Let's say we want to take the top 5-degree nodes and label them.
```{r}
topdeg <- data.frame("deg" = net %v% 'degree') %>% 
  top_n(5)
net %v% 'labels' <- ifelse((net %v% 'degree') %in% topdeg$deg, 
                           net %v% 'name', '')
```

Now we have a sparse label attribute

```{r}
net %v% 'labels'
```

And so it can make our plotting a bit better
```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree'), color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = net %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black")
```

Now that we've got that down as a base, I want to set this as a function so that we can move through other material a little more smoothly
```{r}
netplot_function <- function(netname){
  p <- ggraph(netname, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = netname %v% 'degree'), 
                  color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = netname %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black") 
  return(p)
}
```

## Community detection

Often we're curious about whether or not there is some pattern within that giant plate of spaghetti. Are some organizations linking with certain others on a more regular basis? We might want to do some community detection approaches.

Broadly, community detection is...

# Communities

Have there been certain communities of research over time?

The one thing I think igraph has going for it is community detection.

```{r}
?igraph::cluster_walktrap() # need to specify steps
```

```{r}
wt <- igraph::cluster_walktrap(g)
table(wt$membership) # 47 communities, where most are 
```

In network there is the equivalent cluster function

```{r}
?sna::equiv.clust()
```

```{r}
ec <- equiv.clust(net)
plot(ec)
length(unique(ec$cluster$height))
nblocks = 30
sbm <- blockmodel(net, ec, k = nblocks)
sbm
membership <- data.frame("block" = sbm$block.membership,
                         "id" = ec$cluster$order)
membership <- membership[order(membership$id),]
net %v% 'membership' <- membership$block
table(net %v% 'membership')
```
We could assign color to that membership... but some of those memberships are really not significant

```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree',
                      color = factor(net %v% 'membership', levels = 1:30))) +
  theme_void() +
  geom_node_text(aes(label = net %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black")  +
  scale_color_viridis_d() +
  labs(color = "Community", size = "Degree")
```

```{r}
table(net %v% 'membership')
sig_blocks <- data.frame('corr' = diag(sbm$block.model),
           'b' = 1:30) %>% 
  filter(!(is.nan(corr)) & corr > .5)

net %v% 'member_grp' <- ifelse(net %v% 'membership' == sig_blocks$b[1], 1, 
                          ifelse(net %v% 'membership' == sig_blocks$b[2], 2, 
                          ifelse(net %v% 'membership' == sig_blocks$b[3], 3,
                          ifelse(net %v% 'membership' == sig_blocks$b[4], 4,
                          ifelse(net %v% 'membership' == sig_blocks$b[5], 5,
                          ifelse(net %v% 'membership' == sig_blocks$b[6], 6,
                          ifelse(net %v% 'membership' == sig_blocks$b[7], 7,
                          ifelse(net %v% 'membership' == sig_blocks$b[8], 8,
                                 9))))))))
```

```{r}
ggraph(net, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net %v% 'degree',
                      color = factor(net %v% 'member_grp'))) +
  theme_void() +
  geom_node_text(aes(label = net %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black")  +
  theme(legend.position = "none") +
  scale_color_manual(values = c(clrs, "grey")) 
```

# Across time?

Often you will want to induce subnetworks -- based on edge attribute
```{r}
# Get the edges for each time period
t1 <- which(net %e% 'before_1980' == T)
t2 <- which(net %e% 'Y1980_1994' == T)
t3 <- which(net %e% 'Y1995_2009' == T)
t4 <- which(net %e% 'Y2010_2024' == T)

# Induce subgraphs based on those edges
net_t1 <- network::get.inducedSubgraph(net, eid = t1)
net_t2 <- network::get.inducedSubgraph(net, eid = t2)
net_t3 <- network::get.inducedSubgraph(net, eid = t3)
net_t4 <- network::get.inducedSubgraph(net, eid = t4)
```

```{r}
# Get the edges for each time period
t1_edges <- filter(edges_p, before_1980 == T)
t2_edges <- filter(edges_p, Y1980_1994 == T)
t3_edges <- filter(edges_p, Y1995_2009 == T)
t4_edges <- filter(edges_p, Y2010_2024 == T)
```


Now we have subgraphs
```{r}
net_t4
```

But note that our attributes remain from the whole network
```{r}
(net_t1 %v% 'degree')[(net_t1 %v% 'name' == "SFEI")]
(net_t2 %v% 'degree')[(net_t2 %v% 'name' == "SFEI")]
```

But this is uncorrect, so we want to reassign those values
```{r}
net_t1 %v% 'degree' <- sna::degree(net_t1)
net_t2 %v% 'degree' <- sna::degree(net_t2)
net_t3 %v% 'degree' <- sna::degree(net_t3)
net_t4 %v% 'degree' <- sna::degree(net_t4)
df <- data.frame('degree' = net_t4 %v% 'degree', name = net_t4 %v% 'name')
```

Since our degree value has changed, we also wanted to change our labels. We can write a function to do that

```{r}
get_top_degree <- function(netname, n){
  topdeg <- data.frame("deg" = netname %v% 'degree') %>% 
    top_n(n)
  labels <- ifelse((netname %v% 'degree') %in% topdeg$deg, 
                             netname %v% 'name', '')
  return(labels)
}

net_t1 %v% 'labels' <- get_top_degree(net_t1, 5)
net_t2 %v% 'labels' <- get_top_degree(net_t2, 5)
net_t3 %v% 'labels' <- get_top_degree(net_t3, 5)
net_t4 %v% 'labels' <- get_top_degree(net_t4, 5)
```

Now let's plot one of our subgraphs. We can use the same netplot function
```{r}
netplot_function(net_t1)
netplot_function(net_t2)
netplot_function(net_t3)
netplot_function(net_t4)
```
The issue here is that it is hard to really visualize change because the layout changes every time to make the visualize look more like a network... So let's fixing our coordinates. We want nodes to be in the same position in the plotting of every subgraph.


# Fixing coordinates

With ggraph/network, we can fix coordinates by creating a layout table from our initial network. First let's make sure we have unique identifiers as that table. These layout tables give x and y coordinates to each 
```{r}
net %v% 'id' <- net %v% 'vertex.names'
fixed_coord <- create_layout(net, layout = 'kk')
```

# Then we can create those same layout tables for each subgraph
```{r}
fixed_coord_t4 <- create_layout(net_t4, layout = 'kk')
# joins don't seem to wrok
#fixed_coord_t4 <- fixed_coord_t4[,-c(1:2)]
#fixed_coord_t4 <- right_join(fixed_coord[,c(1,2,6)], fixed_coord_t4)
coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t4$id, c(1,2,6)]
fixed_coord_t4$x <- coords_we_need$x
fixed_coord_t4$y <- coords_we_need$y

fixed_coord_t4  %>% 
  ggraph() + 
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = degree), 
                  color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = labels),
                   family = "Times",
                   size = 3, 
                   color="black") +
  xlim(c(-9, 7)) +
  ylim(c(-7, 8))
```

```{r}
subplot_viz_fixed <- function(fixed_coords, netname){

  fixed_coord_t <- create_layout(netname, layout = 'kk')
  coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t$id, c(1,2,6)]
  fixed_coord_t$x <- coords_we_need$x
  fixed_coord_t$y <- coords_we_need$y
  
  fixed_coord_t  %>% 
    ggraph() + 
    geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
    geom_node_point(aes(size = degree), 
                    color = clrs[6]) +
    theme_void() +
    theme(legend.position = "none") +
    geom_node_text(aes(label = labels),
                     family = "Times",
                     size = 3, 
                     color="black") +
    xlim(c(-9, 7)) +
    ylim(c(-7, 8))
}
```

```{r}
subplot_viz_fixed(fixed_coord, net)
subplot_viz_fixed(fixed_coord, net_t1)
subplot_viz_fixed(fixed_coord, net_t2)
subplot_viz_fixed(fixed_coord, net_t3)
subplot_viz_fixed(fixed_coord, net_t4)
```

```{r}
subplot_viz_fixed(fixed_coord, net) + 
  labs(title = "Collaborative network: 1950-2022") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net_t1) + 
  labs(title = "Collaborative network: 1950-1979") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net_t2) + 
  labs(title = "Collaborative network: 1980-1994") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net_t3) + 
  labs(title = "Collaborative network: 1995-2009") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net_t4) + 
  labs(title = "Collaborative network: 2010-2022") +
  theme(text= element_text(size=10, family="Times")) 
```




```{r}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               label.size = 2,
               legend.position = "none")
```

```{r}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               palette = viridisLite::cividis(n=12),
               label.size = 2,
               legend.position = "none")
```

```{r}
colors <- viridis::viridis(nblocks)
net %v% 'member_color' <- ifelse(net %v% 'membership' %in% c(4,7,9,10,12), colors[7],
                          ifelse(net %v% 'membership' == 1, colors[2], 
                          ifelse(net %v% 'membership' == 2, colors[3], 
                          ifelse(net %v% 'membership' == 3, colors[4], 
                          ifelse(net %v% 'membership' == 5, colors[5], 
                          ifelse(net %v% 'membership' == 6, colors[6], 
                          ifelse(net %v% 'membership' == 8, colors[8], 
                          ifelse(net %v% 'membership' == 11, colors[11], 
                                 colors[12]))))))))
```


```{r}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'member_color',
               label.size = 2,
               legend.position = "none")
```

## Interactive plots
```{r}
plotly::ggplotly(p1)
```

# D3

```{r}
GGally::ggnet2(net,
               size = "degree")
```

```{r}
GGally::ggnet2(net,
               size = "degree",
               color = clrs[6],
               alpha = .5,
               legend.position = "none")
```

```{r}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = clrs[6],
               alpha = .9,
               label.size = 2,
               legend.position = "none")
```