# One-mode visualization

```{r, setup, include=FALSE}
knitr::opts_knit$set(warning = F, message = F, fig.align = 'center')
```

```{r}
library(dplyr)
library(magrittr)
library(igraph)
library(network)
library(sna)
library(ggraph)
```

```{r, echo = F}
edges_1mode <- read.csv('data/edgelist_projected.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')

net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)
g1 <- igraph::graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

## Introduction  

### Visualization goals:  

1. How have different organizations collaborated across projects?  
2. Do we see any particular research communities emerge across these collaborations? (community detection + representing node attributes by color)
3. How have those collaborations changed over time? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)

### Preparing to work with `network`  

We are going to stick to only working with the `network` package, even though the functions we will use for visualization are compatible with both objects. I propose using network objects because they are compatible with more advanced statistical analysis provided through the `statnet` suite of packages. `igraph` objects should also work with the `ggraph` function, but because the syntax for working with the objects is different, the code used to do that would need to change.  

Before we move forward with network, it is worth mentioning that there are commonly-used network functions in each package, `igraph` and `network`, that mask one another. For instance:

```{r, eval = F}
sna::degree()
igraph::degree()
```

The `igraph` degree function will work only on `igraph` objects and `sna` `degree` function will work only on network objects. This can create a headache. To avoid confusion and masking in R, we are going to detach the `igraph` package and work only with `network` objects and compatible packages like `sna`.  

```{r, eval = F}
detach("package:igraph", unload = TRUE)
```

## Getting started with `ggraph`

To introduce the basics of the `ggraph` package, we're going to focus on our first question: _What does the collaborative network of Delta science research look like?_ 

As notes in the previous section, `ggraph` uses the same approach as the `ggplot2` grammar of graphics. This means is has three core components: 1. (network) data, 2. geometries, in this case edges and nodes, and 3. aesthetic mappings. As with any ggplot we can add a theme that suits the look of the figure, and 'void' themes are often suited for networks.

```{r}
ggraph(net1) +
    geom_edge_link() +
    geom_node_point() +
    theme_void()
```

These are some defaults at play here:  

* layout = 'auto', which is the 'stress' layout  
* default aesthetic mapping for geom_edges: x, y, xend, yend
* default aesthetic mapping for geom_node_point: x, y

### Layouts

First, let's talk about layouts. Networks are typically laid out based on the algorithm you (or the package default) chooses. These layouts then are a _choice_, based on how you would like to present the network. Because networks are depicting actors and relationships in an two-dimensional space (x,y), there is no 'right' way to plot actors in space (unless, of course, you have actual spatial data associated with your nodes). In which case, you could use an approach like this NETWORK MAP. Instead, many network theorists have developed layout algorithms, which define rules for calculating the x and y coordinates of the nodes. Many layout algorithms aim to represent networks so that strongly connected nodes are plotting in close proximity.

There is an overview of some of the different layouts presented in this [blog post ](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/), and descriptions of some of them in the [ggraph vignette](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html) and [documentation](https://cran.r-project.org/web/packages/ggraph/ggraph.pdf), under the `layout_tbl_graph_...` functions. Below we present our network with a few different layouts.  

```{r, out.width= c("50%", "50%"), fig.show="hold"}
ggraph(net1, layout = 'stress', bbox = 5) +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Stress (bbox = 5)') +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'kk') +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Kamada and Kawai (spring-based algorithm)')  +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'fr') +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Fruchterman and Reingold (force-directed algorithm)')  +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'circle') +
    geom_edge_link() +
    geom_node_point() +
    theme_void()  +
    labs(title = 'Circle')  +
    theme(plot.title = element_text(hjust = 0.5))
```

Note that there is a whole world of plotting networks using [different conceptualizations of nodes](https://cran.r-project.org/web/packages/ggraph/vignettes/Nodes.html) and [edges](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html). These conceptualizations of nodes and edges may relate to the For the sake of this workshop we are sticking to more 'traditional' network visualization in that we will stay firmly rooted in nodes depicted as points and edges depicted as line.

```{r, echo = F, eval = F}
ggraph(net1, layout = 'tree') +
    geom_edge_diagonal() +
    theme_void() +
    labs(title = 'Tree')
```

You can also set a **manual layout**. Because layouts are just x and y coordinates of points defined by a certain algorithm. You can extract those coordinates using the `create_layout()` function from `ggraph`, and manipulate those coordinates, if need be. We will deal with this more shortly.  

```{r}
fixed_coord <- create_layout(net1, layout = 'fr')
head(fixed_coord[c(1:3,5)])
```

For our one-mode networks, we are going to use the 'fr' layout, which is the Fruchterman and Reingold (force-directed algorithm). This is a choice that helps place higher-degree nodes in the center, and low-degree and isolate nodes on the periphery. 

### Aesthetics

Now that we've chosen a layout, let's start adding some aesthetic features. This process will look very much like aesthetic mapping in `ggplot2`: we can assign aesthetics like color, size, shape, etc. to both our edge and node geometries. We can assign those features from the data directly, or just as stand-alone features.  

**Size by degree**: Let's first say that we'd like to size our nodes by their degree centrality. While more central nodes are already places at the center of our layout's algorithm, it can be helpful to also increase their size to communicate this point. To size by degree, we will want to create a degree variable are a node attribute using the `degree` function. 

```{r}
net1 %v% 'degree' <- sna::degree(net1)
```

**Color-blind friendly colors**: We'd also like our nodes and edges to be colored differently than the default black, so we can set these colors to our geometries. Because the focus of these networks are on nodes, not edges, we can set our edges to a less pronounced color like grey, and select a color-blind friendly color palette to select a node color from. I personally like to use viridis:

```{r}
viridis::viridis(12)
```

The latest version of R studio lets us see these colors when we write them out in a script, so let's do that, and assign these colors to an object named `clrs`. We will be referencing this vector as we start using the palette.

```{r}
clrs <- c("#440154FF", "#482173FF", "#433E85FF", "#38598CFF",
          "#2D708EFF", "#25858EFF", "#1E9B8AFF", "#2BB07FFF", 
          "#51C56AFF", "#85D54AFF", "#C2DF23FF", "#FDE725FF")
```

We can now integrate these three features: node size, edge color, and node color, into our plot. Let's also add an alpha for the edges there.

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70") +
  geom_node_point(aes(size = net1 %v% 'degree'), color = clrs[4]) 
```

Note: There is also some [compatibility with tidyverse's graph function, tidygraph](https://cran.r-project.org/web/packages/ggraph/vignettes/tidygraph.html), but we will not really be exploring that today. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70") +
  geom_node_point(aes(size = tidygraph::centrality_degree()), color = clrs[4]) 
```

We can also add themes and labels, just like with ggplot2. `theme_void()` is great for networks because it removes any coordinates or background color. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70", alpha = .5) +
  geom_node_point(aes(size = net1 %v% 'degree'), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Delta Science Collaborative Research Network")
```
Notice it changes every time, a little bit. You can set your seed (every time before you plot to keep it consistent.)

### Node labels 

So far we are getting a clear shape of the network. But our main goal is to understand _who_ is central to collaboration. Let's try to add some node text with `geom_node_text()`. 

We have a variable that is the name of our vertices:

```{r}
head(net1 %v% 'name')
```

Let's add this as a text geometry.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70", alpha = 0.5) +
  geom_node_point(aes(size = net1 %v% 'degree'), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = net1 %v% 'name'),
                size = 3, 
                color="black")
```

Okay, a bit overwhelming. Instead, let's be selective based on degree. Let's say we want to take the top 5-degree nodes and label them.

```{r}
topdeg <- data.frame("deg" = net1 %v% 'degree') %>% 
  top_n(5)
net1 %v% 'labels' <- ifelse((net1 %v% 'degree') %in% topdeg$deg, 
                           net1 %v% 'name', '')
```

Now we have a sparse label attribute.  

```{r}
net1 %v% 'labels'
```

We can use this labels aesthetic to make our figure more easily readable.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70", alpha = 0.5) +
  geom_node_point(aes(size = net1 %v% 'degree'), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
    labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = net1 %v% 'labels'),
                   size = 3, 
                   color="black")
```

Almost. Let's include a repel = T argument to make sure the text doesn't overlap. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70", alpha = 0.5) +
  geom_node_point(aes(size = net1 %v% 'degree'), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
    labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = net1 %v% 'labels'),
                   size = 3, 
                   color="black",
                 repel = T)
```

Now that we've got that down as a base, I want to set this as a function so that we can move through other material a little more smoothly.  

```{r}
netplot_function <- function(netname){
  p <- ggraph(netname, layout = 'fr') +
  geom_edge_link(color = "gray70", alpha = 0.5) +
  geom_node_point(aes(size = netname %v% 'degree'), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = netname %v% 'labels'),
                   size = 3, 
                   color="black", 
                 repel = T) 
  return(p)
}
```

### Removing isolates 

Before we move too far along in our formatting of this figure, we may want to remove isolates. There are certainly occasions where we _want_ to see isolates in our network, but other times we are interested in the main component. We can identify our isolates and then induce our subgraph with only the non-isolate nodes using the `get.inducedSubgraph()` function.  

```{r}
isolates(net1)
noiso <-  (1:network.size(net1))[-isolates(net1)]
net1_noiso <- get.inducedSubgraph(net1, noiso)
```

Now we can see our network without isolates.  

```{r}
netplot_function(net1_noiso)
```

## Community detection

Moving on to a deeper look at the network. Often we're curious about whether or not there is some pattern within our network. Are some organizations linking with certain others on a more regular basis? Is there a grouping trend in our network that we can't see without some kind of visual aid? To answer this, we might want to do some community detection approaches.

### Community is broadly defined  

Broadly, community detection is...

Have there been certain communities of research over time?

While we've moved away from `igraph`, I do want to point out that it has a number of community detection algorithms. 

```{r, eval = F}
?igraph::cluster_walktrap() 
?igraph::cluster_leiden()
?igraph::cluster_optimal()
```

In the `sna` package there is the equivalence cluster (`equiv.cluster`) function.  

```{r, eval = F}
?sna::equiv.clust()
```

### Equivalence clusters  

Equivalence clustering is a clustering method which groups nodes based on distance from one another. You can specify what kinds of clustering method to use with the `cluster.method` argument and what distance function to use with `equiv.fun`, but it by default uses a hierarchical clustering method based on structural equivalence using Hamming distance. We can calculate the distance and plot a dendrogram of the clusters.  
```{r}
ec <- equiv.clust(net1_noiso)
plot(ec)
```

We can look at the different 'heights' of the nodes within the dendrogram, and see how many different levels there are.  
```{r}
length(unique(ec$cluster$height))
```

Now that we have these clusters, we can turn to another method, blockmodelling, which will evaluate blocks of groupings of nodes. We can specify how may blocks we want (`k`), or specify a height at which to split the classes (`h`). The block model output is a matrix with density measures (this is the default of the function). 
```{r}
bm <- blockmodel(net1_noiso, ec, k = 30,  glabels = "DSC Collaboration")
bm$block.model
```


```{r}
membership <- data.frame("block" = bm$block.membership,
                         "id" = ec$cluster$order)
membership <- membership[order(membership$id),]
net1_noiso %v% 'membership' <- membership$block
table(net1_noiso %v% 'membership')
```

We could assign color to that membership... but some of those memberships may be more or less meaningful. 

```{r}
ggraph(net1_noiso, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net1_noiso %v% 'degree',
                      color = factor(net1_noiso %v% 'membership'))) +
  theme_void() +
  geom_node_text(aes(label = net1_noiso %v% 'labels'),
                   size = 3, 
                   color="black",
                 repel = T)  +
  scale_color_viridis_d() +
  labs(color = "Community", size = "Degree")
```

One way to address this is to consider a threshold for the density values of each block to identify blocks that are more dense than the whole network (see [Slide 92](https://pauloserodio.com/eui2019/day2.pdf)). The density of our network is:

```{r}
gden(net1_noiso)
```

So we can use this as a threshold.   
```{r}
dense_blocks <- data.frame('density' = diag(bm$block.model),
           'b' = 1:length(unique(net1_noiso %v% 'membership'))) %>% 
  filter(!(is.nan(density)) & density > gden(net1_noiso))

net1_noiso %v% 'member_grp' <- ifelse(net1_noiso %v% 'membership' == dense_blocks$b[1], 1, 
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[2], 2, 
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[3], 3,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[4], 4,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[5], 5,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[6], 6,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[7], 7,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[8], 8,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[9], 9,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[10], 10,
                          ifelse(net1_noiso %v% 'membership' == dense_blocks$b[11], 11,
                                 12)))))))))))
```

Now we can use the 'member_grp' node attribute, which associated large and correlated blocks with colors, as our color aesthetic. What we can see is that the most active, central actors like USGS, USFWS, etc., are not in highly dense blocks. Instead, they must work in low density, broader networks in which they span multiple, less connected groups. 

```{r}
ggraph(net1_noiso, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net1_noiso %v% 'degree',
                      color = factor(net1_noiso %v% 'member_grp'))) +
  theme_void() +
  geom_node_text(aes(label = net1_noiso %v% 'labels'),
                   size = 3, 
                   color="black",
                 repel = T)  +
  theme(legend.position = "none") +
  scale_color_manual(values = c(clrs[1:11], "grey")) 
```

## Longitudinal networks  

Often you will want to induce subnetworks -- based on edge attribute

### Induce subgraphs by edge  
```{r}
net1 %v% 'id' <- net1 %v% 'vertex.names'

# Get the edges for each time period
t1 <- which(net1 %e% 'before_1980' == T)
t2 <- which(net1 %e% 'Y1980_1994' == T)
t3 <- which(net1 %e% 'Y1995_2009' == T)
t4 <- which(net1 %e% 'Y2010_2024' == T)

# Induce subgraphs based on those edges
net1_t1 <- network::get.inducedSubgraph(net1, eid = t1)
net1_t2 <- network::get.inducedSubgraph(net1, eid = t2)
net1_t3 <- network::get.inducedSubgraph(net1, eid = t3)
net1_t4 <- network::get.inducedSubgraph(net1, eid = t4)
```

```{r, echo = F, eval = F}
# Get the edges for each time period -- this is just a check
t1_edges <- filter(edges_1mode, before_1980 == T)
t2_edges <- filter(edges_1mode, Y1980_1994 == T)
t3_edges <- filter(edges_1mode, Y1995_2009 == T)
t4_edges <- filter(edges_1mode, Y2010_2024 == T)
```

Now we have subgraphs
```{r}
net1_t4
```

But note that our attributes remain from the whole network
```{r}
(net1_t1 %v% 'degree')[(net1_t1 %v% 'name' == "SFEI")]
(net1_t2 %v% 'degree')[(net1_t2 %v% 'name' == "SFEI")]
```

But this is incorrect, so we want to reassign those values
```{r}
net1_t1 %v% 'degree' <- sna::degree(net1_t1)
net1_t2 %v% 'degree' <- sna::degree(net1_t2)
net1_t3 %v% 'degree' <- sna::degree(net1_t3)
net1_t4 %v% 'degree' <- sna::degree(net1_t4)
```

Since our degree value has changed, we also wanted to change our labels. We can write a function to do that

```{r}
get_top_degree <- function(netname, n){
  topdeg <- data.frame("deg" = netname %v% 'degree') %>% 
    top_n(n)
  labels <- ifelse((netname %v% 'degree') %in% topdeg$deg, 
                             netname %v% 'name', '')
  return(labels)
}

net1_t1 %v% 'labels' <- get_top_degree(net1_t1, 5)
net1_t2 %v% 'labels' <- get_top_degree(net1_t2, 5)
net1_t3 %v% 'labels' <- get_top_degree(net1_t3, 5)
net1_t4 %v% 'labels' <- get_top_degree(net1_t4, 5)
```

Now let's plot one of our subgraphs. We can use the same netplot function
```{r, out.width = c("50%", "50%")}
netplot_function(net1_t1)
netplot_function(net1_t2)
netplot_function(net1_t3)
netplot_function(net1_t4)
```
The issue here is that it is hard to really visualize change because the layout changes every time to make the visualize look more like a network... So let's fixing our coordinates. We want nodes to be in the same position in the plotting of every subgraph.


### Fixing coordinates

With ggraph/network, we can fix coordinates by creating a layout table from our initial network. First let's make sure we have unique identifiers as that table. These layout tables give x and y coordinates to each 
```{r}
fixed_coord <- create_layout(net1, layout = 'fr')
```

With these coordinates fixed from our full plot, we can then apply those same coordinates for each subgraph.

```{r}
# create a layout table for the subgraph
fixed_coord_t1 <- create_layout(net1_t1, layout = 'fr')
# subset the relevant coordinates from the full layout table
coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t1$id, c(1,2,6)]
fixed_coord_t1$x <- coords_we_need$x
fixed_coord_t1$y <- coords_we_need$y

fixed_coord_t1  %>% 
  ggraph() + 
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = degree), 
                  color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = labels),
                   family = "Times",
                   size = 3, 
                   color="black") +
  xlim(c(-9, 7)) +
  ylim(c(-7, 8))
```

```{r}
subplot_viz_fixed <- function(fixed_coords, netname){

  fixed_coord_t <- create_layout(netname, layout = 'fr')
  coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t$id, c(1,2,6)]
  fixed_coord_t$x <- coords_we_need$x
  fixed_coord_t$y <- coords_we_need$y
  
  fixed_coord_t  %>% 
    ggraph() + 
    geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
    geom_node_point(aes(size = degree), 
                    color = clrs[4]) +
    theme_void() +
    theme(legend.position = "none") +
    geom_node_text(aes(label = labels),
                     family = "Times",
                     size = 3, 
                     color="black") +
    xlim(c(-9, 7)) +
    ylim(c(-7, 8))
}
```

```{r}
subplot_viz_fixed(fixed_coord, net1)
subplot_viz_fixed(fixed_coord, net1_t1)
subplot_viz_fixed(fixed_coord, net1_t2)
subplot_viz_fixed(fixed_coord, net1_t3)
subplot_viz_fixed(fixed_coord, net1_t4)
```

```{r}
subplot_viz_fixed(fixed_coord, net1) + 
  labs(title = "Collaborative network: 1950-2022") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t1) + 
  labs(title = "Collaborative network: 1950-1979") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t2) + 
  labs(title = "Collaborative network: 1980-1994") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t3) + 
  labs(title = "Collaborative network: 1995-2009") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t4) + 
  labs(title = "Collaborative network: 2010-2022") +
  theme(text= element_text(size=10, family="Times")) 
```

## Interactive plots DRAFTING
```{r}
p1 <- subplot_viz_fixed(fixed_coord, net1) + 
  labs(title = "Collaborative network: 1950-2022") +
  theme(text= element_text(size=10, family="Times")) 
plotly::ggplotly(p1)
```

D3


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               label.size = 2,
               legend.position = "none")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               palette = viridisLite::cividis(n=12),
               label.size = 2,
               legend.position = "none")
```

```{r, eval = F, echo = F}
colors <- viridis::viridis(nblocks)
net %v% 'member_color' <- ifelse(net %v% 'membership' %in% c(4,7,9,10,12), colors[7],
                          ifelse(net %v% 'membership' == 1, colors[2], 
                          ifelse(net %v% 'membership' == 2, colors[3], 
                          ifelse(net %v% 'membership' == 3, colors[4], 
                          ifelse(net %v% 'membership' == 5, colors[5], 
                          ifelse(net %v% 'membership' == 6, colors[6], 
                          ifelse(net %v% 'membership' == 8, colors[8], 
                          ifelse(net %v% 'membership' == 11, colors[11], 
                                 colors[12]))))))))
```


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'member_color',
               label.size = 2,
               legend.position = "none")
```


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               color = clrs[6],
               alpha = .5,
               legend.position = "none")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = clrs[6],
               alpha = .9,
               label.size = 2,
               legend.position = "none")
```