# One-mode visualization

```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, fig.align = 'center')
```

```{r, warning = F, message = F, echo = F}
library(igraph)
library(network)
library(ggraph)
```

```{r, echo = F}
edges_1mode <- read.csv('data/edgelist_projected.csv')
nodes_1mode <- read.csv('data/nodelist_onemode.csv')

net1 <- network(x = edges_1mode, 
               vertices = nodes_1mode,
               bipartite = F,  
               directed = F)
g1 <- igraph::graph_from_data_frame(edges_1mode, vertices = nodes_1mode, directed = F)
```

## Introduction  

### Preparing to work with `network`  

For the remainder of this tutorial are going to stick to working with the `network` package, even though the functions we will use for visualization are compatible with both objects. I propose using network objects because they are compatible with more advanced statistical analysis provided through the `statnet` suite of packages. `igraph` objects should also work with the `ggraph` function, but because the syntax for working with the objects is different, the code used to do that would need to change.  

To continue working with `network` objects and `ggraph`, we want to load in two more packages. The `sna` package provides us with several functions for calculating network statistics, and `dplyr` and `magrittr` are packages from that tidyverse that will help us streamline our plotting.  

```{r}
library(sna)
library(dplyr)
library(magrittr)
```

Because we will be using exclusively `network` objects, we want to detach `igraph` before we continue further. This is because there are several commonly-used network functions in both `igraph` and `sna` that mask one another. For instance:

```{r, eval = F}
sna::degree()
igraph::degree()
```

The `igraph` degree function will work only on `igraph` objects and `sna` `degree` function will work only on `network` objects. To avoid confusion and masking in R, we are going to detach the `igraph` package and work only with `network` objects and compatible packages like `sna`.  

```{r}
detach("package:igraph", unload = TRUE)
```

### Guiding questions:  

To get started with visualization, let's remind ourselves of our visualization goals and guiding questions. With our one-mode network, we want to know: 

1. What is the structure of the collaborative research network in the Delta?
2. Do we see any particular research communities emerge across these collaborations? (community detection + representing node attributes by color)
3. How have those collaborations changed over time? (subgraphs based on edge attribute + multiple plots with fixed coordinates)

## Getting started with `ggraph`

To introduce the basics of the `ggraph` package, we're going to focus on our first question: _What is the structure of the collaborative research network in the Delta?_ 

As described in the previous section, `ggraph` uses the same approach as the `ggplot2` grammar of graphics. This means is has three core components: 1.) (network) data, 2.) geometries, in this case edges and nodes, and 3.) aesthetic mappings. As with any ggplot we can add a theme that suits the look of the figure, and 'void' themes are often suited for networks.

```{r, message = T}
ggraph(net1) +
    geom_node_point() +
    geom_edge_link() +
    theme_void()
```

These are some defaults at play here:  

* The layout argument is set to = 'auto', which is the 'stress' layout  
* The default aesthetic mapping for geom_node_point: x, y, which don't need specified  
* The default aesthetic mapping for geom_edges: x, y, xend, yend, which don't need specified  

### Layouts

First, let's talk about layouts. Networks are typically laid out based on the algorithm you (or the package default) chooses. Layouts are a _choice_ based on how you would like to present the network. Because networks are depicting actors and relationships in an two-dimensional space (x, y), there is no 'right' way to plot actors in space (unless, of course, you have actual spatial data associated with your nodes). In which case, you could use a geospatial network mapping (e.g. [flight route maps](https://news.delta.com/route-map-us-canada)). For scenarios without geospatial coordinates, many network theorists have developed layout algorithms which define rules for calculating the x and y coordinates of the nodes. Many layout algorithms aim to represent networks so that strongly connected nodes are plotted in close proximity to one another, representing a 'core' and a 'periphery'.  

There is an overview of `ggraph`'s different layouts presented in this [blog post ](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/), and descriptions of some of them in the [ggraph vignette](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html) and [documentation](https://cran.r-project.org/web/packages/ggraph/ggraph.pdf), under the `layout_tbl_graph_...` functions. The layout is selected with the 'layout' argument within the ggraph function. Below we present our network with a few different layouts.  

```{r, out.width= c("50%", "50%"), fig.show="hold", echo = F}
ggraph(net1, layout = 'stress', bbox = 5) +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Stress: `layout = "stress", bbox = 5`') +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'kk') +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Kamada and Kawai (spring-based algorithm): `layout = "kk"`')  +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'fr') +
    geom_edge_link() +
    geom_node_point() +
    theme_void() +
    labs(title = 'Fruchterman and Reingold (force-directed algorithm): `layout = "fr"`')  +
    theme(plot.title = element_text(hjust = 0.5))

ggraph(net1, layout = 'circle') +
    geom_edge_link() +
    geom_node_point() +
    theme_void()  +
    labs(title = 'Circle: `layout = "circle"`')  +
    theme(plot.title = element_text(hjust = 0.5))
```

Note that there is a whole world of plotting networks using [different conceptualizations of nodes](https://cran.r-project.org/web/packages/ggraph/vignettes/Nodes.html) and [edges](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html). These conceptualizations of nodes and edges may be dependent on certain layout algorithms. For example, if working with a 'tree' based layout, which creates more of a hierarchical structure like a dendrogram, edges as lines may not be suited. Instead, you can specify different edges (e.g. diagonals).  

```{r, echo = T,  out.width = "50%", fig.align='center'}
ggraph(net1, layout = 'tree') +
    geom_edge_diagonal() +
    theme_void() +
    labs(title = 'Tree: `layout = "tree"`') +
    theme(plot.title = element_text(hjust = 0.5))
```


For the sake of this workshop we are sticking to more 'traditional' network visualization, so we will only be depicting nodes as points and edges as lines.  
 
You can also set a **manual layout**. Because layouts are just x and y coordinates of points defined by a certain algorithm, you can extract those coordinates using the `create_layout()` function from `ggraph`, and fix and/or manipulate those coordinates, if need be. We will deal with this more shortly.  

```{r}
fixed_coord <- create_layout(net1, layout = 'fr')
head(fixed_coord[c(1:2,5)])
```

For our one-mode networks, we are going to use the 'fr' layout, which is the Fruchterman and Reingold (force-directed algorithm). This is a choice that helps place higher-degree nodes in the center, and low-degree and isolate nodes on the periphery. 

```{r}
ggraph(net1, layout = 'fr') +
    geom_node_point() +
    geom_edge_link() +
    theme_void()
```

### Aesthetics

Now that we've chosen a layout, let's start adding some aesthetic features. This process will look very much like aesthetic mapping in `ggplot2`: we can assign aesthetics like color, size, shape, etc. to both our edge and node geometries. 

**Size by degree**: One common approach for network visualization is to size nodes by their degree centrality. While more central nodes are already places at the center of our layout's algorithm, it can be helpful to also increase their size to communicate this point. To size by degree, we will want to create a degree variable as a node attribute using the `degree` function, and assign that attribute to our network data.  

```{r}
net1 %v% 'degree' <- degree(net1)
```

Hint: if you get `Error in degree(net1) : Not a graph object`, double check that you have detached the `igraph` package!  

**Color-blind friendly colors**: We'd also like our nodes and edges to be colored differently than the default black, so we can set these colors to our geometries. Because the focus of these networks are on nodes, not edges, we can set our edges to a less pronounced color like grey, and select an [accessible, color-blind friendly color palette](https://ucdavisdatalab.github.io/workshop_data_viz_principles/accessible-data-visualizations.html) to select a node color from. I personally like to use viridis:

```{r}
viridis::viridis(12)
```

The latest version of R studio lets us see these colors when we write them out in a script, so let's do that, and assign these colors to an object named `clrs`. We will be referencing this vector as we start using the palette.

```{r}
clrs <- c("#440154FF", "#482173FF", "#433E85FF", "#38598CFF",
          "#2D708EFF", "#25858EFF", "#1E9B8AFF", "#2BB07FFF", 
          "#51C56AFF", "#85D54AFF", "#C2DF23FF", "#FDE725FF")
```

We can now integrate these three features: node size, edge color, and node color, into our plot. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_node_point(aes(size = degree), 
                  color = clrs[4]) +
  geom_edge_link(color = "gray80") +
  theme_void()
```

Notice how layer order matters (as as with `ggplot2`), and so by having edges layered on top of nodes, we are really hiding the nodes. Let's try to switch this, but also include some alpha arguments to help increase transparency.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(size = degree), 
                  color = clrs[4], alpha = .7) +
  theme_void()
```

<!----
Note: The approach we will take for this workshop is to assign attributes to the network data frame, then add them into our plotting aesthetics. There is also some [compatibility with tidyverse's graph package, tidygraph](https://cran.r-project.org/web/packages/ggraph/vignettes/tidygraph.html), which we could use to set network aesthetics directly in the plotting functions (below), but we will not really be exploring that today. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray70") +
  geom_node_point(aes(size = tidygraph::centrality_degree()), 
                  color = clrs[4], alpha = .8) +
  theme_void()
```
---->

We can also add labels and make thematic alterations to features like the legend, just like with ggplot2.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(size = degree), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  labs(title = "Delta Science Collaborative Research Network") +
  theme(legend.position = "none")
```
Notice that with the 'fr' layout (and any other layout algorithm), the coordinates change a bit every time. This is because each time we create a visualization the algorithm is re-run, and there is variation in the exact calculation. You can set your seed (every time before you plot) to keep it consistent)

### Node labels 

So far we are getting a clear shape of the network. But related to our first question about understanding network structure, we may want to understand  _who_ is central to collaboration. To better identify our nodes, let's try to add some node text with `geom_node_text()`. Already, we have a variable that is the name of our vertices:

```{r}
head(net1 %v% 'name')
```

Let's add this as a text geometry.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(size = degree), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = name),
                size = 3, 
                color="black")
```

Okay, a bit overwhelming. Instead, let's be selective based on degree. Let's say we want to take the top 5-degree nodes and label them.

```{r}
# Use extract the network's degree values based on the order of degrees
degs <- (net1 %v% 'degree')[order(net1 %v% 'degree')]
# Then identify the top 5 unique degree values
topdegs <- unique(rev(degs))[1:5]
# Then create a network variable named labels and add the name only if a
# node has the number of degrees in the 'top degrees'
net1 %v% 'labels' <- ifelse((net1 %v% 'degree') %in% topdegs, 
                           net1 %v% 'name', '')
```

Now we have a sparse label attribute.  

```{r}
net1 %v% 'labels'
```

We can use this labels attribute to make our figure more easily readable.  

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(size = degree), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
    labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = labels),
                   size = 3)
```

Almost. Let's include a repel = T argument to make sure the text doesn't overlap. 

```{r}
ggraph(net1, layout = 'fr') +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(size = degree), color = clrs[4],
                  alpha = .7) +
  theme_void() +
  theme(legend.position = "none") +
    labs(title = "Delta Science Collaborative Research Network") +
  geom_node_text(aes(label = labels),
                   size = 3,
                   repel = T)
```

So, _what is the structure of the collaborative research network in the Delta?_ This network shows us the overall picture, and  we can use summary statistics to fill in the gaps. Among the 139 organizations involved in scientific research in the Delta, there is a mean degree of 14, suggesting that organizations are involved with, on average, 14 other organizations (across one or more projects). The network is quite connected, as the main component includes 107 (77%) of the organizations, with 27 isolates, meaning that 27 organizations have not collaborated at all. Within that main component, the average path length is 2.3, meaning that on average an organization is less then 3 connections away from any other organization. At the center of the network are three federal agencies, the [US Geological Survey (USGS)](https://sciencetracker.deltacouncil.ca.gov/node/49743), [US Fish and Wildlife Service (USFWS)](https://sciencetracker.deltacouncil.ca.gov/node/49741), and [US Bureau of Reclamation (USBR)](https://sciencetracker.deltacouncil.ca.gov/node/49731), two state agencies, [California Department of Fish and Wildlife (CDFW)](https://sciencetracker.deltacouncil.ca.gov/node/49602) and [California Department of Water Resources (DWR)](https://sciencetracker.deltacouncil.ca.gov/node/49609) one research institute, the [San Francisco Estuary Institute (SFEI)](https://sciencetracker.deltacouncil.ca.gov/node/49707). 

<details>
<summary>**Reveal:** How to calculate network-level statistics with the `sna` package </summary>
```{r, purl = T}
network.size(net1)
mean(net1 %v% 'degree')
sum(component.largest(net1))
length(isolates(net1))
main_comp <- component.largest(net1, result = 'graph')
mean(geodist(main_comp)[['gdist']], na.rm = T)
```
</details>
<br>

Now that we've got that down as a base, I want to create this network visualization approach as a function so that we can move through other material a little more smoothly. Feel free to just copy this function -- all we are doing is taking the code we previously wrote, and replacing the network name that we've been using, net1, with the generic argument for the network name, netname. 

```{r}
netplot_function <- function(netname){
  p <- ggraph(netname, layout = 'fr') +
         geom_edge_link(color = "gray80") +
         geom_node_point(aes(size = degree), color = clrs[4],
                         alpha = .7) +
         theme_void() +
         theme(legend.position = "none") +
           labs(title = "Delta Science Collaborative Research Network") +
         geom_node_text(aes(label = labels),
                          size = 3,
                          repel = T)
  return(p)
}
```

### Removing isolates 

Before we move too far along in our formatting of this figure, we may want to remove isolates. There are certainly occasions where we _want_ to see isolates in our network, but other times we are interested in the main component. We can identify our isolates and then induce our subgraph with only the non-isolate nodes using the `get.inducedSubgraph()` function.  

```{r}
isolates(net1)
noiso <-  (1:network.size(net1))[-isolates(net1)]
net1_noiso <- get.inducedSubgraph(net1, noiso)
```

Now we can see our network without isolates, and quickly use our new `netplot_function`:  

```{r}
netplot_function(net1_noiso)
```

## Community detection

Now, let's combine some network statistics for visualization to take a deeper look at the network structure and address our second guiding question: _Do we see any particular research communities emerge across these collaborations?_. Often researchers are curious about whether or not there is some pattern within the network. Though the majority of the nodes are connected in one relatively dense component, are some organizations linking with others on a more regular basis? Is there a grouping trend in our network that we can't see from the network figure on its own? To answer this, we might want to use some community detection approaches.

### Community is broadly defined  

Broadly, community detection is an approach that employs a wide range of distance/clustering algorithms to identify to find connected communities within a network based on features like density/modularity. While we've moved away from `igraph`, I do want to point out that it has a number of community detection algorithms. You can read more about community detection algorithms [here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6103523/). 

```{r, eval = F}
?igraph::cluster_walktrap() 
?igraph::cluster_label_prop()
?igraph::cluster_fast_greedy
```

In the `sna` package there is the equivalence cluster (`equiv.cluster`) function, which identifies distances based on structural equivalence. The resource above, Hoffman et al. 2018, define it as follows: 'Structural equivalence is an important concept from the blockmodeling literature. It is defined where two nodes play the exact same role in a network in terms of their edges, and would be considered interchangeable (Doreian et al., 2005).'  

```{r, eval = F}
?sna::equiv.clust()
```

### Equivalence clusters  

Equivalence clustering is a clustering method which groups nodes based on distance from one another. You can specify what kinds of clustering method to use with the `cluster.method` argument and what distance function to use with `equiv.fun`, but it by default uses a hierarchical clustering method based on structural equivalence using Hamming distance. We can calculate the distance and plot a dendrogram of the clusters.  

```{r}
ec <- equiv.clust(net1_noiso)
plot(ec)
```
<!---
We can look at the different 'heights' of the nodes within the dendrogram, and see how many different levels there are.  
```{r}
length(unique(ec$cluster$height))
```
---> 

Now that we have these clusters, we can turn to another method, blockmodelling, which will evaluate the density of node groupings (or 'blocks') estimated by the equivalence clusters. We can specify how may blocks we want (`k`), or specify a height at which to split the classes (`h`).  We will pick a rather arbitrary number for blocks, k = 5. 

```{r}
bm <- blockmodel(net1_noiso, ec, k = 5,  glabels = "DSC Collaboration")
```

The block model output includes a vector assigning each node to a block through 'block membership'. We can wrap this output in the `table` function to get a sense of how many nodes have been combined across the different blocks. Most of the nodes (97) are in block 1, while the remaning blocks are much smaller, and block 5 has only one node.  
```{r}
table(bm$block.membership)
```

Additionally, the block model itself provides a matrix with a density measure for the within and between block pairings. Looking along the diagonal, we see that the first block has very low density (0.03), but the smaller blocks have high within-block density (1, .89, 1), while there is no value for block 5 because the density cannot be calculated for a network of 1. [CHECK: IS THIS WHAT'S HAPPENING HERE?]

```{r}
bm$block.model
```
We can then assign the block membership to the network. It is worth noting, however, that the vector output of `bm$block.membership` does not maintain the same node order as the nodes in our network. So we want to make sure we reorder the membership vector to align with the vertex names, which are listed under `bm$plabels`. 

```{r}
membership <- data.frame("block" = bm$block.membership,
                         "vertex.names" = bm$plabels)
membership <- membership[order(membership$vertex.names),]
net1_noiso %v% 'membership' <- membership$block
```

```{r}
ggraph(net1_noiso, layout = 'fr') +
         geom_edge_link(color = "gray80") +
         geom_node_point(aes(size = degree, color = factor(membership)),
                         alpha = .7) +
         theme_void() +
         labs(title = "Delta Science Collaborative Research Network") +
         geom_node_text(aes(label = labels),
                          size = 3,
                          repel = T) +
         scale_color_viridis_d() +
         labs(color = "Community", size = "Degree")
```

Let's rethink our labels to get a better sense of who is in what group
```{r}
# Like before, set a conditional statement so that we only have labels
# for nodes whose membership is in the dense nodes
net1_noiso %v% 'member_label' <- ifelse(net1_noiso %v% 'membership' %in% 2:5,
                                        net1_noiso %v% 'name', "")
```

We might also want to make a table of those so we can check it out
```{r}
data.frame('block' = 
             (net1_noiso %v% 'membership')[net1_noiso %v% 'membership' %in% 2:5],
           'name' = (net1_noiso %v% 'name')[net1_noiso %v% 'membership' %in% 2:5]) %>% 
  arrange(block)
```

<!--- additional resource: (see [Slide 92](https://pauloserodio.com/eui2019/day2.pdf)).--->

```{r, warning = T}
iso_net <- ggraph(net1_noiso, layout = 'fr') +
         geom_edge_link(color = "gray80") +
         geom_node_point(aes(size = degree, color = factor(membership)),
                         alpha = .7) +
         theme_void() +
         labs(title = "Delta Science Collaborative Research Network") +
         geom_node_text(aes(label = member_label),
                          size = 3,
                          repel = T) +
         scale_color_viridis_d() +
         labs(color = "Community", size = "Degree")
iso_net
```

And what if I want to clean this up, say, have only one legend and move the legend below. We can specify which legends we want with `guides()` and put the legend in the 'bottom' position with `themes()`.

```{r, warning = T}
iso_net +
    labs(color = "Community", size = "Degree") + 
    guides(color = "legend", size = "none") +
    theme(legend.position = "bottom")
```

Do we see any particular research communities emerge across these collaborations?

To conclude this section we can revisit the second guiding question: _Do we see any particular research communities emerge across these collaborations?_ We've used the `blockmodel` function, paired with more conditional assignments of network attributes to identify some possible research clusters. Among our five 'blocks' we see high density among three of them: blocks 2, 3, and 4. Block 2 includes three of the most central nodes -- CDFW, USBR, USFWS -- suggesting that these three wild-life and land use agencies at the state and federal levels repeatedly collaborate together in the Delta. Block 3 includes a wider variety of state agencies that are closely clustered. And block 4 has only three federal agencies, USDA, EPA, and NOAA, who much like block 2 seem to be consistently collaborating on science projects. Block 5 includes only USGS, the network's most central node, pointing to the challenge of clustering highly central nodes given that they span so many connections. The remainder of the organizations are grouped into Block 1, a low-density mass that have no clear trend towards clustering.  

## Longitudinal networks  

Often you will want to induce subnetworks -- based on edge attribute

### Induce subgraphs by edge  
```{r}
net1 %v% 'id' <- net1 %v% 'vertex.names'

# Get the edges for each time period
t1 <- which(net1 %e% 'before_1980' == T)
t2 <- which(net1 %e% 'Y1980_1994' == T)
t3 <- which(net1 %e% 'Y1995_2009' == T)
t4 <- which(net1 %e% 'Y2010_2024' == T)

# Induce subgraphs based on those edges
net1_t1 <- network::get.inducedSubgraph(net1, eid = t1)
net1_t2 <- network::get.inducedSubgraph(net1, eid = t2)
net1_t3 <- network::get.inducedSubgraph(net1, eid = t3)
net1_t4 <- network::get.inducedSubgraph(net1, eid = t4)
```

```{r, echo = F, eval = F}
# Get the edges for each time period -- this is just a check
t1_edges <- filter(edges_1mode, before_1980 == T)
t2_edges <- filter(edges_1mode, Y1980_1994 == T)
t3_edges <- filter(edges_1mode, Y1995_2009 == T)
t4_edges <- filter(edges_1mode, Y2010_2024 == T)
```

Now we have subgraphs
```{r}
net1_t4
```

But note that our attributes remain from the whole network
```{r}
(net1_t1 %v% 'degree')[(net1_t1 %v% 'name' == "SFEI")]
(net1_t2 %v% 'degree')[(net1_t2 %v% 'name' == "SFEI")]
```

But this is incorrect, so we want to reassign those values
```{r}
net1_t1 %v% 'degree' <- sna::degree(net1_t1)
net1_t2 %v% 'degree' <- sna::degree(net1_t2)
net1_t3 %v% 'degree' <- sna::degree(net1_t3)
net1_t4 %v% 'degree' <- sna::degree(net1_t4)
```

Since our degree value has changed, we also wanted to change our labels. We can write a function to do that

```{r}
get_top_degree <- function(netname, n){
  topdeg <- data.frame("deg" = netname %v% 'degree') %>% 
    top_n(n)
  labels <- ifelse((netname %v% 'degree') %in% topdeg$deg, 
                             netname %v% 'name', '')
  return(labels)
}

net1_t1 %v% 'labels' <- get_top_degree(net1_t1, 5)
net1_t2 %v% 'labels' <- get_top_degree(net1_t2, 5)
net1_t3 %v% 'labels' <- get_top_degree(net1_t3, 5)
net1_t4 %v% 'labels' <- get_top_degree(net1_t4, 5)
```

Now let's plot one of our subgraphs. We can use the same netplot function
```{r, out.width = c("50%", "50%")}
netplot_function(net1_t1)
netplot_function(net1_t2)
netplot_function(net1_t3)
netplot_function(net1_t4)
```
The issue here is that it is hard to really visualize change because the layout changes every time to make the visualize look more like a network... So let's fixing our coordinates. We want nodes to be in the same position in the plotting of every subgraph.


### Fixing coordinates

With ggraph/network, we can fix coordinates by creating a layout table from our initial network. First let's make sure we have unique identifiers as that table. These layout tables give x and y coordinates to each 
```{r}
fixed_coord <- create_layout(net1, layout = 'fr')
```

With these coordinates fixed from our full plot, we can then apply those same coordinates for each subgraph.

```{r}
# create a layout table for the subgraph
fixed_coord_t1 <- create_layout(net1_t1, layout = 'fr')
# subset the relevant coordinates from the full layout table
coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t1$id, c(1,2,6)]
fixed_coord_t1$x <- coords_we_need$x
fixed_coord_t1$y <- coords_we_need$y

fixed_coord_t1  %>% 
  ggraph() + 
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = degree), 
                  color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = labels),
                   family = "Times",
                   size = 3, 
                   color="black") +
  xlim(c(-9, 7)) +
  ylim(c(-7, 8))
```

```{r}
subplot_viz_fixed <- function(fixed_coords, netname){

  fixed_coord_t <- create_layout(netname, layout = 'fr')
  coords_we_need <- fixed_coord[fixed_coord$id %in% fixed_coord_t$id, c(1,2,6)]
  fixed_coord_t$x <- coords_we_need$x
  fixed_coord_t$y <- coords_we_need$y
  
  fixed_coord_t  %>% 
    ggraph() + 
    geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
    geom_node_point(aes(size = degree), 
                    color = clrs[4]) +
    theme_void() +
    theme(legend.position = "none") +
    geom_node_text(aes(label = labels),
                     family = "Times",
                     size = 3, 
                     color="black") +
    xlim(c(-9, 7)) +
    ylim(c(-7, 8))
}
```

```{r}
subplot_viz_fixed(fixed_coord, net1)
subplot_viz_fixed(fixed_coord, net1_t1)
subplot_viz_fixed(fixed_coord, net1_t2)
subplot_viz_fixed(fixed_coord, net1_t3)
subplot_viz_fixed(fixed_coord, net1_t4)
```

```{r}
subplot_viz_fixed(fixed_coord, net1) + 
  labs(title = "Collaborative network: 1950-2022") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t1) + 
  labs(title = "Collaborative network: 1950-1979") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t2) + 
  labs(title = "Collaborative network: 1980-1994") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t3) + 
  labs(title = "Collaborative network: 1995-2009") +
  theme(text= element_text(size=10, family="Times")) 
subplot_viz_fixed(fixed_coord, net1_t4) + 
  labs(title = "Collaborative network: 2010-2022") +
  theme(text= element_text(size=10, family="Times")) 
```

## Interactive plots DRAFTING
```{r}
p1 <- subplot_viz_fixed(fixed_coord, net1) + 
  labs(title = "Collaborative network: 1950-2022") +
  theme(text= element_text(size=10, family="Times")) 
plotly::ggplotly(p1)
```

D3


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               label.size = 2,
               legend.position = "none")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'membership',
               palette = viridisLite::cividis(n=12),
               label.size = 2,
               legend.position = "none")
```

```{r, eval = F, echo = F}
colors <- viridis::viridis(nblocks)
net %v% 'member_color' <- ifelse(net %v% 'membership' %in% c(4,7,9,10,12), colors[7],
                          ifelse(net %v% 'membership' == 1, colors[2], 
                          ifelse(net %v% 'membership' == 2, colors[3], 
                          ifelse(net %v% 'membership' == 3, colors[4], 
                          ifelse(net %v% 'membership' == 5, colors[5], 
                          ifelse(net %v% 'membership' == 6, colors[6], 
                          ifelse(net %v% 'membership' == 8, colors[8], 
                          ifelse(net %v% 'membership' == 11, colors[11], 
                                 colors[12]))))))))
```


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = 'member_color',
               label.size = 2,
               legend.position = "none")
```


```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               color = clrs[6],
               alpha = .5,
               legend.position = "none")
```

```{r, eval = F, echo = F}
GGally::ggnet2(net,
               size = "degree",
               label = 'labels',
               color = clrs[6],
               alpha = .9,
               label.size = 2,
               legend.position = "none")
```