# Two-mode visualization

* What actors have been most active across projects? 
* Are most-active members typically leaders or contributors? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)
* What actors have been most active on climate-related projects? (Node attribute of project node)

We can use the same approaches as with one mode... but we shouldn't because mode one and two are just not comparable.

```{r, echo = F}
library(sna)
library(network)
library(magrittr)
library(ggraph)
library(dplyr)
edges_2mode <- read.csv('data/edgelist_twomode.csv')
nodes_2mode <- read.csv('data/nodelist_twomode.csv')
net2 <- network(x = edges_2mode, 
               vertices = nodes_2mode[nodes_2mode$id %in% edges_2mode$org_id |
                                nodes_2mode$id %in% edges_2mode$project_id,],
               bipartite = T,  
               directed = F)
clrs <- c("#440154FF", "#46337EFF", "#365C8DFF", "#277F8EFF",
          "#1FA187FF", "#4AC16DFF", "#9FDA3AFF", "#FDE725FF")
```

```{r, echo = F}
netplot_function <- function(netname){
  p <- ggraph(netname, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = netname %v% 'degree'), 
                  color = clrs[6]) +
  theme_void() +
  theme(legend.position = "none") +
  geom_node_text(aes(label = netname %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black", repel = T) 
  return(p)
}
```

```{r}
net2 %v% 'degree' <- sna::degree(net2)
topdeg <- data.frame("deg" = net2 %v% 'degree') %>% 
  top_n(5)
net2 %v% 'labels' <- ifelse((net2 %v% 'degree') %in% topdeg$deg, 
                           net2 %v% 'name', '')
netplot_function(net2)
```

```{r}
net2 %v% 'mode'
```

```{r}
ggraph(net2, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net2 %v% 'degree',
                      shape = factor(net2 %v% 'mode'),
                      color = factor(net2 %v% 'mode'))) +
  theme_void() +
  scale_color_viridis_d() +
  geom_node_text(aes(label = net2 %v% 'labels'),
                   size = 3, 
                   color="black") +
    labs(size = "Degree", color = "Mode", shape = "Mode")
```

# Normalized degree for both modes... (see Ebba)

```{r}
# NORMALIZED degrees and label
net2 %v% 'degree_norm' <- sna::degree(net2, rescale = T)

# make a table, project id as rows and org id as columns
#inc_mat <- table(edges_2mode[c('project_id', 'org_id')])
degree_norm_bip <- bipartite::ND(as.matrix(net2))
# output has two measures of degree: one for 'lower' and one for 'higher'
# 'lower' is for for the rows, so in our case, projects
length(degree_norm_bip$lower) == length(unique(edges_2mode$project_id))
# this is compatible with mode 0, which are the second half of the nodes
net2 %v% 'mode'
```

```{r}
# So we want to assign this bipartite degree in accordance with that order
# BUT, note that this function orders degree based on numeric order of id,
# which may not align with the node order
nd_output <- c(degree_norm_bip$higher, degree_norm_bip$lower)
net2 %v% 'vertex.names' == names(nd_output)

nd_output_df <- data.frame('degreeN' = nd_output,
                           'id' = names(nd_output))
nd_output_sorted <- nd_output_df[match((net2 %v% 'vertex.names'), 
                                        nd_output_df $id),]

net2 %v% 'vertex.names' == nd_output_sorted$id
net2 %v% 'degree_norm_bip' <- nd_output_sorted$degreeN

# Note: normalizing the degree for the whole network vs. normalizing degree for
# each mode is a different process
(net2 %v% 'degree_norm')[net2 %v% 'vertex.names' == 49590]
(net2 %v% 'degree_norm_bip')[net2 %v% 'vertex.names' == 49590]
```

```{r}
topdeg_mode0 <- data.frame("deg" = (net2 %v% 'degree_norm_bip')[net2 %v% 'mode' == 0]) %>% 
  top_n(3)
topdeg_mode1 <- data.frame("deg" = (net2 %v% 'degree_norm_bip')[net2 %v% 'mode' == 1]) %>% 
  top_n(3)


net2 %v% 'labels_bip' <- ifelse((net2 %v% 'degree_norm_bip') %in% topdeg_mode0$deg & net2 %v% 'mode' == 0, net2 %v% 'name', 
                                ifelse((net2 %v% 'degree_norm_bip') %in% topdeg_mode1$deg & net2 %v% 'mode' == 1, net2 %v% 'name', ''))
```

```{r}
ggraph(net2, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net2 %v% 'degree_norm_bip',
                      shape = factor(net2 %v% 'mode'),
                      color = factor(net2 %v% 'mode'))) +
  theme_void() +
  scale_color_viridis_d() +
  geom_node_text(aes(label = net2 %v% 'labels_bip'),
                   size = 2, 
                   color="black") +
    labs(size = "Degree", color = "Mode", shape = "Mode")
```

But I want to demonstrate that there can be no connections between projects or between collaborators, in this network, so let's pick a layout that demonstrates that

```{r}
net2 %v% 'type' <- ifelse(net2 %v% 'mode' == 0, T, F)
ggraph(net2, layout = 'bipartite') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net2 %v% 'degree_norm_bip',
                      shape = factor(net2 %v% 'mode'),
                      color = factor(net2 %v% 'mode')),
                  alpha = .75) +
  theme_void() +
  scale_color_viridis_d() +
  geom_node_text(aes(label = net2 %v% 'labels_bip'),
                   size = 2, 
                   color="black") +
    labs(size = "Degree", color = "Mode", shape = "Mode")
```

Add a guide: I only want one legend
```{r}
ggraph(net2, layout = 'bipartite') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net2 %v% 'degree_norm_bip',
                      shape = factor(net2 %v% 'mode'),
                      color = factor(net2 %v% 'mode')),
                  alpha = .75) +
  theme_void() +
  scale_color_viridis_d() +
  geom_node_text(aes(label = net2 %v% 'labels_bip'),
                   size = 2, 
                   color="black") +
    labs(color = "Mode", shape = "Mode") + 
  guides(colour = "legend", shape = "legend", size = "none")
```


## Leaders versus contributors?  

We can return to the idea of subnetworks.

### Induce subgraphs by edge  
```{r}
# Induce subgraphs based on those edges
net2_l <- network::get.inducedSubgraph(net2, 
                                       eid = which(net2 %e% 'leadership' == T))
net2_c <- network::get.inducedSubgraph(net2, 
                                       eid = which(net2 %e% 'contribution' == T))
```

Now we have subgraphs
```{r}
net2_l
```

## Climate projects

```{r}
climate <- which(net2 %v% 'sci_Weatherandclimate' == T)
landuse <- which(net2 %v% 'sci_Landuseandhumanactivities' == T)
govnc <- which(net2 %v% 'mgmt_Governance' == T)
flood <- which(net2 %v% 'mgmt_Floodriskandlandusemanagement' == T)

# this is harder for bipartite
net2_cli <- network::get.inducedSubgraph(net2, 
                                       v = climate)
```

```{r}
net2_cli
```

```{r}
climate_alter_ids <- edges_2mode$org_id[edges_2mode$project_id %in% 
                     (net2 %v% 'vertex.names')[climate]]
landuse_alter_ids <- edges_2mode$org_id[edges_2mode$project_id %in% 
                     (net2 %v% 'vertex.names')[landuse]]
govnc_alter_ids <- edges_2mode$org_id[edges_2mode$project_id %in% 
                     (net2 %v% 'vertex.names')[govnc]]
flood_alter_ids <- edges_2mode$org_id[edges_2mode$project_id %in% 
                     (net2 %v% 'vertex.names')[flood]]
```

```{r}
climate_alters <- which(net2 %v% 'vertex.names' %in% climate_alter_ids)
landuse_alters <- which(net2 %v% 'vertex.names' %in% landuse_alter_ids)
govnc_alters <- which(net2 %v% 'vertex.names' %in% govnc_alter_ids)
flood_alters <- which(net2 %v% 'vertex.names' %in% flood_alter_ids)

net2_cli <- network::get.inducedSubgraph(net2, 
                                       v = climate,
                                       alters = climate_alters)
net2_landuse <- network::get.inducedSubgraph(net2, 
                                       v = landuse,
                                       alters = landuse_alters)
net2_govnc <- network::get.inducedSubgraph(net2, 
                                       v = govnc,
                                       alters = govnc_alters)
net2_flood <- network::get.inducedSubgraph(net2, 
                                       v = flood,
                                       alters = flood_alters)
```


```{r}

assign_bipartite_deg_labels <- function(netname, n){
  degree_norm_bip <- bipartite::ND(as.matrix(netname))
  nd_output <- c(degree_norm_bip$higher, degree_norm_bip$lower)
  nd_output_df <- data.frame('degreeN' = nd_output,
                             'id' = names(nd_output))
  nd_output_sorted <- nd_output_df[match((netname %v% 'vertex.names'), 
                                          nd_output_df $id),]
  
  netname %v% 'degree_norm_bip' <- nd_output_sorted$degreeN
  
  topdeg_mode0 <- data.frame("deg" = 
              (netname %v% 'degree_norm_bip')[netname %v% 'mode' == 0]) %>%
    top_n(n)
  topdeg_mode1 <- data.frame("deg" = 
              (netname %v% 'degree_norm_bip')[netname %v% 'mode' ==   1]) %>% 
    top_n(n)

  netname %v% 'labels_org' <- ifelse((netname %v% 'degree_norm_bip') %in%
                                       topdeg_mode1$deg & 
                                       netname %v% 'mode' == 1, 
                                     netname %v% 'name', '')
  return(netname)
}

```

```{r}
net2_cli <- assign_bipartite_deg_labels(net2_cli, 3)
net2_landuse <- assign_bipartite_deg_labels(net2_landuse, 3)
net2_govnc <- assign_bipartite_deg_labels(net2_govnc, 3)
net2_flood <- assign_bipartite_deg_labels(net2_flood, 3)
```

```{r}
plot_bipartite_fx <- function(netname){
  p <- ggraph(netname, layout = 'bipartite') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = netname %v% 'degree_norm_bip',
                      shape = factor(netname %v% 'mode'),
                      color = factor(netname %v% 'mode')),
                  alpha = .75) +
  theme_void() +
  scale_color_viridis_d() +
  geom_node_text(aes(label = netname %v% 'labels_org'),
                   size = 2, 
                   color="black") +
    labs(color = "Mode", shape = "Mode") + 
  guides(colour = "legend", shape = "legend", size = "none")
  return(p)
}
```

```{r}
plot_bipartite_fx(net2_cli)
plot_bipartite_fx(net2_landuse)
plot_bipartite_fx(net2_govnc)
plot_bipartite_fx(net2_flood)
```

