# Two-mode visualization

* What actors have been most active across projects? 
* Are most-active members typically leaders or contributors? (make subgraphs based on edge attribute + make multiple plots with fixed coordinates)
* What actors have been most active on climate-related projects? (Node attribute of project node)

We can use the same approaches as with one mode... but we shouldn't because mode one and two are just not comparable.

```{r, echo = F}
edges <- read.csv('data/edgelist_twomode.csv')
nodes <- read.csv('data/nodelist_twomode.csv')
net_b <- network(x = edges, 
               vertices = nodes[nodes$id %in% edges$org_id |
                                nodes$id %in% edges$project_id,],
               bipartite = T,  
               directed = F)
```

```{r}
net_b %v% 'degree' <- sna::degree(net_b)
topdeg <- data.frame("deg" = net_b %v% 'degree') %>% 
  top_n(5)
net_b %v% 'labels' <- ifelse((net_b %v% 'degree') %in% topdeg$deg, 
                           net_b %v% 'name', '')
netplot_function(net_b)
```

```{r}
net_b %v% 'mode'
```

```{r}
ggraph(net_b, layout = 'fr') +
  geom_edge_link(width = 0.5, alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = net_b %v% 'degree',
                      shape = factor(net_b %v% 'mode')), 
                  color = clrs[6]) +
  theme_void() +
  geom_node_text(aes(label = net_b %v% 'labels'),
                   family = "Times",
                   size = 3, 
                   color="black")
```

# Normalized degree for both modes... (see Ebba)

# Introductio to `ggbipart`
